---
title: "Drug Group Cosinor Thoracic IL-6 Functionalized"
output: html_document
date: "2024-12-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# List of packages to check and install
packages <- c("base", "cosinor", "cosinor2", "ggplot2", "grid", "gridExtra",
        "here", "magrittr", "minpack.lm", "openxlsx", "readxl", "stats", "tidyr", "dplyr")

# Function to check and install missing packages
missing_packages <- packages[!packages %in% installed.packages()[, "Package"]]
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

# Load all packages
lapply(packages, library, character.only = TRUE)

library(cosinor)
library(ggplot2)
library(gridExtra)
library(here)
library(minpack.lm)
library(readxl)
library(tidyr)
library(dplyr)
```


```{r}
# Sourcing Custom-Made Functions
source("../outlier_detection.R")
source("../filter_outliers.R")
source("../extract_cosinor.R")
source("../zero_amp.R")
source("../plot_cosinor.R")
source("../create_combined_pdf.R")
source("../export_excel.R")
```


### Reading, Converting, Analyzing and Plotting Thoracic IL-6 
### Administered Rats 
#### Reading in Thoracic IL-6 Administered Rat Data for 
#### LD and rLEN
```{r}
# Reading Thoracic IL-6 Administered Rats for LD
readxl::read_excel(path ="../Summary Multiplex Data.xlsx",
                   sheet = "Thoracic",
                    range ="N38:R44") -> LD_Circadian_Thor_IL6_Filter

# Reading Thoracic IL-6 Administered Rats for rLEN
readxl::read_excel(path ="../Summary Multiplex Data.xlsx",
                   sheet = "Thoracic",
                    range ="U38:Y44") -> rLEN_Circadian_Thor_IL6_Filter
```


#### Converting Data into Usable Format for Extracting VOIs and Plotting
```{r}
# Data Conversion to get Individual Points for LD Thoracic IL-6 
# Rat Model
## X value is Zeitgeber time, Y value is Concentration in mg of Drug/Content

# Loads Initial Dataset for LD Thoracic IL-6 Rats and 
# Removes Outliers
#outlier_detection(Circadian_Thor_IL6_LD) -> out_res_LDTIL6
#filter_outliers(Circadian_Thor_IL6_LD, 
#                out_res_LDTIL6) ->  LD_Circadian_Thor_IL6_Filter

# Performs Data Carpentry on Filtered LD IL6 Dataset and Removes 'NA's,
## filters out ZT of 24
LD_Circadian_Thor_IL6_Filter %>%
  pivot_longer(cols = c(`0`,`6`,`12`, `18`, `24`), names_to = "ZT", 
               values_to = "Concentration(mg)") %>%
  drop_na(`Concentration(mg)`) %>%
  mutate(ZT = as.numeric(ZT)) %>%
  arrange(ZT) -> LD_Circadian_Thor_IL6_Filter
LD_Circadian_Thor_IL6_Filter %>%
  # Filter out ZT of 24
  filter(ZT != 24)  -> LD_Circadian_Thor_IL6_data

# Creating Cosinor Model for LD IL6 Rat Plots
cosinor.lm(`Concentration(mg)` ~ time(ZT), period = 24,
           data = LD_Circadian_Thor_IL6_Filter) -> cosMod_LD_TIL6
# Creating Cosinor Model for LD IL6 Rat Data Analysis
cosinor.lm(`Concentration(mg)` ~ time(ZT), period = 24,
           data = LD_Circadian_Thor_IL6_data) -> cosMod_LD_TIL6_data
```

```{r}
# Data Conversion to get Individual Points for rLEN Thoracic IL-6 
# Rat Model X value is Zeitgeber time, Y value is Concentration in 
# mg of Drug/Content

# Loads Initial Dataset for rLEN Thoracic IL-6 Rats and Removes Outliers
#outlier_detection(Circadian_Thor_IL6_rLEN) -> out_res_rLENTIL6
#filter_outliers(Circadian_Thor_IL6_rLEN, 
#                out_res_rLENTIL6) ->  rLEN_Circadian_Thor_IL6_Filter

# Performs Data Carpentry on Filtered LD IL6 Dataset and Removes 'NA's
rLEN_Circadian_Thor_IL6_Filter %>%
  pivot_longer(cols = c(`0`,`6`,`12`, `18`, `24`), names_to = "ZT", 
               values_to = "Concentration(mg)") %>%
  drop_na(`Concentration(mg)`) %>%
  mutate(ZT = as.numeric(ZT)) %>%
  arrange(ZT) -> rLEN_Circadian_Thor_IL6_Filter
rLEN_Circadian_Thor_IL6_Filter %>%
  # Filter out ZT of 24
  filter(ZT != 24)  -> rLEN_Circadian_Thor_IL6_data

# Creating Cosinor Model for LD IL6 Rat Plots
cosinor.lm(`Concentration(mg)` ~ time(ZT), period = 24,
           data = rLEN_Circadian_Thor_IL6_Filter) -> cosMod_rLEN_TIL6
# Creating Cosinor Model for LD IL6 Rat Plot Data Analysis
cosinor.lm(`Concentration(mg)` ~ time(ZT), period = 24,
           data = rLEN_Circadian_Thor_IL6_data) -> cosMod_rLEN_TIL6_data
```


#### Calculating the Y-limits for Plots of Cosinor Fit
```{r}
# Finding the max Y-value Across Both Datsets in Order to Set Y-limits of 
# Future Plot
max(rLEN_Circadian_Thor_IL6_Filter$`Concentration(mg)`, 
    LD_Circadian_Thor_IL6_Filter$`Concentration(mg)`) -> TIL6_max
ceiling(TIL6_max / 5) * 5 -> y_peak
```


#### Creating Plots of Original Data with Cosinor Fit Overlayed for Rat Models
```{r}
# Creating Individual Plots of Rat Datasets w/ Cosinor Fits
## Loads Function plot_cosinor Into Global Environment, Which Takes cosMod 
## Model, Original Dataset, Plot Title and Desired x-breaks (currently set for
## Zetigeber Time, to see how to change to regular time see function notes) and
## Generates Plot of Original Data with Overlayed Cosinor Fit Model

# Creating Plot for LD IL6 Rat Model
plot_cosinor(cosMod_LD_TIL6, LD_Circadian_Thor_IL6_Filter, 
             "LD Thoracic IL-6", "LD",
             x_breaks = seq(0, 24, by = 6)) -> mod_plot_LD_TIL6


# Creating Plot for rLEN IL6 Rat Model
plot_cosinor(cosMod_rLEN_TIL6, rLEN_Circadian_Thor_IL6_Filter, 
             "rLEN Thoracic IL-6", "rLEN",
             x_breaks = seq(0, 24, by = 6)) -> mod_plot_rLEN_TIL6
```


#### Performing Cosinor Analysis for Rat Datasets
```{r}
# Creating Table of Data VOIs for LD Thoracic IL-6 Rat Model
## Produces Table of Pertinent Values (Mesor, Amplitude, Acrophase, Bathyphase)
## and Converts Table into Graphical Object for Figure Creation
extract_cosinor(cosMod_LD_TIL6_data, LD_Circadian_Thor_IL6_data,
                "LD") -> LD_extract_output
LD_TIL6_Results_table <- LD_extract_output$results_table
LD_TIL6_Results_grob <- LD_extract_output$results_grob


# Creating Table of Data VOIs for rLEN Thoracic IL-6 Rat Model
## Produces Table of Pertinent Values (Mesor, Amplitude, Acrophase, Bathyphase)
## and Converts Table into Graphical Object for Figure Creation
extract_cosinor(cosMod_rLEN_TIL6_data, rLEN_Circadian_Thor_IL6_data,
                "rLEN") -> rLEN_extract_output
rLEN_TIL6_Results_table <- rLEN_extract_output$results_table
rLEN_TIL6_Results_grob <- rLEN_extract_output$results_grob
```


#### Performing Zero Amplitude Test Analysis on Rat Models
```{r}
# Zero-Amplitude Analysis of Cosinor Fit
## Creating Table of Zero Amplitude VOIs for LD Thoracic IL-6 Rat Model

# Produces Table of Pertinent Values (df1, df2, f-value, p-value) and Converts 
# Table into Graphical Object for Figure Creation
zero_amp(cosMod_LD_TIL6_data, "LD") -> LD_zero_output
LD_TIL6_amp_test_Results <- LD_zero_output$results_table
LD_TIL6_amp_test_grob <- LD_zero_output$results_grob


# Produces Table of Pertinent Values (df1, df2, f-value, p-value) and Converts 
# Table into Graphical Object for Figure Creation
zero_amp(cosMod_rLEN_TIL6_data, "rLEN") -> rLEN_zero_output
rLEN_TIL6_amp_test_Results <- rLEN_zero_output$results_table
rLEN_TIL6_amp_test_grob <- rLEN_zero_output$results_grob
```


#### Creating Side-by-Side Plot Comparison of Cosinor Fit Models for LD and 
#### rLEN Rat Models
```{r, fig.width= 15}
# Set the working directory to the folder where you want to save the plot
plot_dir <- here("Combined Plots")
# Create the directory if it does not exist
if (!dir.exists(plot_dir)) dir.create(plot_dir)  

# Save the plot as PDF in the specified directory
pdf(file.path(plot_dir, "combined_TIL6_plots.pdf"), width=15, height=6)
grid.arrange(mod_plot_LD_TIL6, mod_plot_rLEN_TIL6, ncol = 2)
dev.off()
```


#### Creating Side-by-Side Table Comparisons of Cosinor Analysis and Zero
#### Amplitude Analysis VOIs
```{r}
# Set the working directory to the folder where you want to save the tables
table_dir <- here("Combined Tables")
# Create the directory if it does not exist
if (!dir.exists(table_dir)) dir.create(table_dir)  

headers <- c("Cosinor Parameters", "Zero Amp Parameters")
tables <- list(LD_TIL6_Results_grob, LD_TIL6_amp_test_grob, 
               rLEN_TIL6_Results_grob, rLEN_TIL6_amp_test_grob, NULL, NULL)
title_text <- "Comparison of Thoracic IL-6 LD and rLEN Circadian Disruption Values"
output_file <- file.path(table_dir, "combined_TIL6_tables.pdf")
fig_width <- 7.5
fig_height <- 10
create_combined_pdf(headers, tables, title_text, output_file, fig_width, fig_height)
```


#### Exporting all Created Table Figures to Excel for Easse of
#### Data Copying/Compiling
```{r}
# Creating File Marker for Table Figure Identification in Function Code
file_abbrev <- "TIL6"
# Exporting Table Statistics to Excel Sheets
export_excel(file_abbrev)
```