---
title: "Duplicated Group Cosinor Clock Per2 Functionalized"
output: html_document
date: "2025-01-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# List of packages to check and install
packages <- c("base", "cosinor", "cosinor2", "ggplot2", "grid", "gridExtra",
        "here", "magrittr", "minpack.lm", "openxlsx", "readxl", "readr", "stats",
        "tidyr", "dplyr")

# Function to check and install missing packages
missing_packages <- packages[!packages %in% installed.packages()[, "Package"]]
if (length(missing_packages) > 0) {
  install.packages(missing_packages)
}

# Load all packages
lapply(packages, library, character.only = TRUE)

library(cosinor)
library(ggplot2)
library(gridExtra)
library(here)
library(minpack.lm)
library(readxl)
library(readr)
library(tidyr)
library(dplyr)
```


```{r}
# Sourcing Custom-Made Functions
source("../outlier_detection.R")
source("../filter_outliers.R")
source("../extract_cosinor_clock.R")
source("../zero_amp.R")
source("../plot_cosinor_clock.R")
source("../create_combined_pdf.R")
source("../export_excel.R")
```


### Reading, Converting, Analyzing and Plotting Clock Per2 
### Fold Changes in Rats 
#### Reading in Clock Per2 Administered Rat Data for 
#### LD and rLEN
```{r}
# Reading Clock Per2 Administered Rats
readxl::read_excel(path ="../SCN PCR Data Summary.xlsx",
                   sheet = "Per2",
                    range ="A1:M40") ->  Circadian_Clock_Per2

Circadian_Clock_Per2 %>%
  select(`Animal ID`,`Time point`, `Group`, `Normalized fold change`) %>%
  rename("ID" = `Animal ID`, "Time" = `Time point`,
        "Change" = `Normalized fold change`)-> Circadian_Clock_Per2


# Check if 'Normalized fold change' is character and needs parsing
if (is.character(Circadian_Clock_Per2$Change)) {
  Circadian_Clock_Per2 <- Circadian_Clock_Per2 %>%
    mutate(Change = parse_number(Change))
}

# Separating LD data and pivoting
Circadian_Clock_Per2_LD <- Circadian_Clock_Per2%>%
  filter(grepl("LD", Group)) %>%
  mutate(
    Time = recode(Time,
    "ZT0" = "0", "ZT6" = "6", "ZT12" = "12", "ZT18" = "18", "ZT24" = "24")
  ) %>%
  pivot_wider(
    id_cols = ID,
    names_from = Time,
    values_from = Change) %>%
  select(-c(ID))
if (!"24" %in% colnames(Circadian_Clock_Per2_LD)) {
  LD_Circadian_Clock_Per2_Filter<- Circadian_Clock_Per2_LD %>%
    mutate(`24` = `0`)}

# Separating rLEN data and pivoting    
Circadian_Clock_Per2_rLEN <- Circadian_Clock_Per2 %>%
  filter(grepl("rLEN", Group)) %>%
  mutate(
    Time = recode(Time,
    "ZT0" = "0", "ZT6" = "6", "ZT12" = "12", "ZT18" = "18", "ZT24" = "24")
  ) %>%
  pivot_wider(
    id_cols = ID,
    names_from = Time,
    values_from = Change)%>%
  select(-c(ID))
if (!"24" %in% colnames(Circadian_Clock_Per2_rLEN)) {
  rLEN_Circadian_Clock_Per2_Filter <- Circadian_Clock_Per2_rLEN %>%
    mutate(`24` = `0`)}
```


#### Converting Data into Usable Format for Extracting VOIs and Plotting
```{r}
# Data Conversion to get Individual Points for LD Clock Per2 
# Rat Model
## X value is Zeitgeber time, Y value is # of Fold Changes

# Loads Initial Dataset for LD Clock Per2 Rats and 
# Removes Outliers
#outlier_detection(Circadian_Clock_Per2_LD) -> out_res_LDCPer2
#filter_outliers(Circadian_Clock_Per2_LD, 
                #out_res_LDCPer2) ->  LD_Circadian_Clock_Per2_Filter

# Performs Data Carpentry on Filtered LD TNFa Dataset and Removes 'NA's,
## filters out ZT of 24
LD_Circadian_Clock_Per2_Filter %>%
  pivot_longer(cols = c(`0`,`6`,`12`, `18`, `24`), names_to = "ZT", 
               values_to = "Fold_Changes") %>%
  drop_na(`Fold_Changes`) %>%
  mutate(ZT = as.numeric(ZT)) %>%
  arrange(ZT) -> LD_Circadian_Clock_Per2_Filter
#LD_Circadian_Clock_Per2_Filter %>%
  # Filter out ZT of 24
  #filter(ZT != 24)  -> LD_Circadian_Clock_Per2_data

# Creating Cosinor Model for LD TNFa Rat Plots
cosinor.lm(`Fold_Changes` ~ time(ZT), period = 24,
           data = LD_Circadian_Clock_Per2_Filter) -> cosMod_LD_CPer2
# Creating Cosinor Model for LD TNFa Rat Data Analysis
#cosinor.lm(`Fold_Changes` ~ time(ZT), period = 24,
#           data = LD_Circadian_Clock_Per2_data) -> cosMod_LD_CPer2_data
```

```{r}
# Data Conversion to get Individual Points for rLEN Clock Per2 
# Rat Model X value is Zeitgeber time, Y value is # of Fold Changes

# Loads Initial Dataset for rLEN Clock Per2 Rats and Removes Outliers
#outlier_detection(Circadian_Clock_Per2_rLEN) -> out_res_rLENCPer2
#filter_outliers(Circadian_Clock_Per2_rLEN, 
 #               out_res_rLENCPer2) ->  rLEN_Circadian_Clock_Per2_Filter

# Performs Data Carpentry on Filtered LD TNFa Dataset and Removes 'NA's
rLEN_Circadian_Clock_Per2_Filter %>%
  pivot_longer(cols = c(`0`,`6`,`12`, `18`, `24`), names_to = "ZT", 
               values_to = "Fold_Changes") %>%
  drop_na(`Fold_Changes`) %>%
  mutate(ZT = as.numeric(ZT)) %>%
  arrange(ZT) -> rLEN_Circadian_Clock_Per2_Filter
#rLEN_Circadian_Clock_Per2_Filter %>%
  # Filter out ZT of 24
  #filter(ZT != 24)  -> rLEN_Circadian_Clock_Per2_data

# Creating Cosinor Model for LD TNFa Rat Plots
cosinor.lm(`Fold_Changes` ~ time(ZT), period = 24,
           data = rLEN_Circadian_Clock_Per2_Filter) -> cosMod_rLEN_CPer2
# Creating Cosinor Model for LD TNFa Rat Plot Data Analysis
#cosinor.lm(`Fold_Changes` ~ time(ZT), period = 24,
#           data = rLEN_Circadian_Clock_Per2_data) -> cosMod_rLEN_CPer2_data
```


#### Calculating the Y-limits for Plots of Cosinor Fit
```{r}
# Finding the max Y-value Across Both Datsets in Order to Set Y-limits of 
# Future Plot
max(rLEN_Circadian_Clock_Per2_Filter$`Fold_Changes`, 
    LD_Circadian_Clock_Per2_Filter$`Fold_Changes`) -> CPer2_max
# Calculate y_peak based on CPer2_max
if (CPer2_max < 4) {
   y_peak <- ceiling(CPer2_max / 0.25) * 0.25 
  } else {
  y_peak <- ceiling(CPer2_max / 5) * 5}
```


#### Creating Plots of Original Data with Cosinor Fit Overlayed for Rat Models
```{r}
# Creating Individual Plots of Rat Datasets w/ Cosinor Fits
## Loads Function plot_cosinor Into Global Environment, Which Takes cosMod 
## Model, Original Dataset, Plot Title and Desired x-breaks (currently set for
## Zetigeber Time, to see how to change to regular time see function notes) and
## Generates Plot of Original Data with Overlayed Cosinor Fit Model

# Creating Plot for LD TNFa Rat Model
plot_cosinor_clock(cosMod_LD_CPer2, LD_Circadian_Clock_Per2_Filter, 
             "LD Clock Per2", "LD",
             x_breaks = seq(0, 24, by = 6)) -> mod_plot_LD_CPer2


# Creating Plot for rLEN TNFa Rat Model
plot_cosinor_clock(cosMod_rLEN_CPer2, rLEN_Circadian_Clock_Per2_Filter, 
             "rLEN Clock Per2", "rLEN",
             x_breaks = seq(0, 24, by = 6)) -> mod_plot_rLEN_CPer2
```


#### Performing Cosinor Analysis for Rat Datasets
```{r}
# Creating Table of Data VOIs for LD Clock Per2 Rat Model
## Produces Table of Pertinent Values (Mesor, Amplitude, Acrophase, Bathyphase)
## and Converts Table into Graphical Object for Figure Creation
extract_cosinor_clock(cosMod_LD_CPer2, LD_Circadian_Clock_Per2_Filter,
                "LD") -> LD_extract_output
LD_CPer2_Results_table <- LD_extract_output$results_table
LD_CPer2_Results_grob <- LD_extract_output$results_grob


# Creating Table of Data VOIs for rLEN Clock Per2 Rat Model
## Produces Table of Pertinent Values (Mesor, Amplitude, Acrophase, Bathyphase)
## and Converts Table into Graphical Object for Figure Creation
extract_cosinor_clock(cosMod_rLEN_CPer2, rLEN_Circadian_Clock_Per2_Filter,
                "rLEN") -> rLEN_extract_output
rLEN_CPer2_Results_table <- rLEN_extract_output$results_table
rLEN_CPer2_Results_grob <- rLEN_extract_output$results_grob
```


#### Performing Zero Amplitude Test Analysis on Rat Models
```{r}
# Zero-Amplitude Analysis of Cosinor Fit
## Creating Table of Zero Amplitude VOIs for LD Clock Per2 Rat Model

# Produces Table of Pertinent Values (df1, df2, f-value, p-value) and Converts 
# Table into Graphical Object for Figure Creation
zero_amp(cosMod_LD_CPer2, "LD") -> LD_zero_output
LD_CPer2_amp_test_Results <- LD_zero_output$results_table
LD_CPer2_amp_test_grob <- LD_zero_output$results_grob


# Produces Table of Pertinent Values (df1, df2, f-value, p-value) and Converts 
# Table into Graphical Object for Figure Creation
zero_amp(cosMod_rLEN_CPer2, "rLEN") -> rLEN_zero_output
rLEN_CPer2_amp_test_Results <- rLEN_zero_output$results_table
rLEN_CPer2_amp_test_grob <- rLEN_zero_output$results_grob
```


#### Creating Side-by-Side Plot Comparison of Cosinor Fit Models for LD and 
#### rLEN Rat Models
```{r, fig.width= 15}
# Set the working directory to the folder where you want to save the plot
plot_dir <- here("Combined Plots")
# Create the directory if it does not exist
if (!dir.exists(plot_dir)) dir.create(plot_dir)  

# Save the plot as PDF in the specified directory
pdf(file.path(plot_dir, "duplicated_combined_CPer2_plots.pdf"), width=15, height=6)
grid.arrange(mod_plot_LD_CPer2, mod_plot_rLEN_CPer2, ncol = 2)
dev.off()
```


#### Creating Side-by-Side Table Comparisons of Cosinor Analysis and Zero
#### Amplitude Analysis VOIs
```{r}
# Set the working directory to the folder where you want to save the tables
table_dir <- here("Combined Tables")
# Create the directory if it does not exist
if (!dir.exists(table_dir)) dir.create(table_dir)  

headers <- c("Cosinor Parameters", "Zero Amp Parameters")
tables <- list(LD_CPer2_Results_grob, LD_CPer2_amp_test_grob, 
               rLEN_CPer2_Results_grob, rLEN_CPer2_amp_test_grob, NULL, NULL)
title_text <- "Comparison of Clock Per2 LD and rLEN Fold Change Values (w/ Duplicated Values)"
output_file <- file.path(table_dir, "duplicated_combined_CPer2_tables.pdf")
fig_width <- 7.5
fig_height <- 10
create_combined_pdf(headers, tables, title_text, output_file, fig_width, fig_height)
```


#### Exporting all Created Table Figures to Excel for Easse of
#### Data Copying/Compiling
```{r}
# Creating File Marker for Table Figure Identification in Function Code
file_abbrev <- c("CPer2", "duplicated")
# Exporting Table Statistics to Excel Sheets
export_excel(file_abbrev)
```
